---
title: 'Lab 3: Topic Analysis'
author: "Javier Patrón"
output:
  word_document: default
  pdf_document: default
---

```{r packages}
library(quanteda)
library(tm)
library(topicmodels)
library(ldatuning)
library(tidyverse)
library(tidytext)
library(reshape2)

```

Load the data

```{r data}

#tbl <-read_csv("/Users/javipatron/Documents/MEDS/Courses/eds231/text_modeling-lab2/spring_text")
tbl <-read_csv("https://raw.githubusercontent.com/MaRo406/EDS-231-text-sentiment/main/data/sb_dat_text.csv")

```

First we'll build the corpus using corpus() from {quanteda}.

```{r corpus}

corp_sb <- corpus(x = tbl, text_field = "text")
stories_stats <- summary(corp_sb)
head(stories_stats)

```

Next, we can use tokens(), also from {quanteda}, to construct a tokens object. tokens() takes a range of arguments related to cleaning the data. Next we'll create a stopwords lexicon and remove each word contained in it from our tokens object. The quanteda function tokens_select() lets us do the removal.

```{r tokens}
toks <- tokens(corp_sb, remove_punct = T, remove_numbers = T)
add_stops <- c(stopwords("en"), "spring", "break")

toks1 <- tokens_select(toks, pattern = add_stops, selection = "remove")

```

Now we can transform our data to prepare for topic modeling. Let's create a document-feature matrix with quanteda::dfm(). Topic modeling doesn't work with empty rows in this matrix, so we'll need to remove those. I do that here using {slam}, which is designed to deal with sparse matrices like ours.

```{r dfm}


# Create the matrix with the texts
dfm_sb <- dfm(toks1, tolower = T)
dfm <- dfm_trim(dfm_sb, min_docfreq = 2)
sel_idx <- slam::row_sums(dfm) > 0
dfm <- dfm[sel_idx,]

```

Great, now we are almost ready to run a model. We just have to come up with an initial value for k, the number of latent topics present in the data. How do we do this? There are multiple methods. I think there are three main ideas you (well, ChatGPT) wrote about in your stories: weather, travel, and relaxation. So I will tell the model to look for 3 topics by setting the k parameter = 3.

```{r LDA_modeling}

# Set the number of topics you want
k <- 3

topicModel_k3 <- LDA(dfm, k, method = "Gibbs", 
                     control = list(iter = 500),
                     verbose = 25)

```

Running topicmodels::LDA() produces two posterior probability distributions: theta, a distribution over k topics within each document and beta,the distribution v terms within each topic, where v is our vocabulary (total unique words in our data set).

Let's examine at our results. posterior() extracts theta and beta

```{r LDA_modeling}
result <- posterior(topicModel_k3)

attributes(result)

beta <- result$terms
theta <- result$topics

dim(beta)
dim(theta)


terms(topicModel_k3, 10)
dim(terms(topicModel_k3, 20))


```

Alright, so that worked out OK. An alternative to specifying k based on theory or a hypothesis is to run a series of models using a range of k values. ldatuning::FindTopicsNumber gives us the tools for this.

```{r find_k}

result <- FindTopicsNumber(dfm, 
                           topics = seq(from = 2,
                                        to = 20, 
                                        by = 1), 
                           metrics = c("CaoJuan2009", "Deveaud2014"),
                           method = "Gibbs",
                           verbose = T )

FindTopicsNumber_plot(result)

```

Alright, now let's estimate another model, this time with our new value of k.

```{r LDA_again}
k <- 10

topicModel_k10 <- LDA(dfm, k, method="Gibbs", control=list(iter = 500, verbose = 25))

tmResult <- posterior(topicModel_k10)
terms(topicModel_k10, 10)
theta <- tmResult$topics #distributions of each of the topics. Here is where you create and tune your topics for future analysis
beta <- tmResult$terms #distribution of each of the terms
vocab <- (colnames(beta))

```

There are multiple proposed methods for how to measure the best k value. You can [go down the rabbit hole here](https://rpubs.com/siri/ldatuning)

```{r top_terms_topic}

sb_topics <- tidy(topicModel_k10, matrix = "beta")
top_terms <- sb_topics |> 
  group_by(topic) |> 
  top_n(10, beta) |> 
  ungroup() |> 
  arrange(topic, -beta)


top_terms
```

```{r plot_top_terms}

top_terms %>%
  mutate(term = reorder_within(term, beta, topic, sep = "")) %>%
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free_y") +
  scale_x_reordered()+
  coord_flip()

```

Let's assign names to the topics so we know what we are working with. We can name the topics by interpreting the overall theme or idea they represent, but for now let's just name them by their top terms.

```{r topic_names}

topic_words <- terms(topicModel_k10, 5)
topic_names <- apply(topic_words, 2, paste, collapse = " ")

```

We can explore the theta matrix, which contains the distribution of each topic over each document.

```{r topic_dists}

example_ids <- c(1:5)
n <- length(example_ids)

# get topic proportions from example documents
example_props <- theta[example_ids,] # Filter by the first 5 in the example_ids 

colnames(example_props) <- topic_names


viz_df <- melt(cbind(data.frame(example_props),
                     document = factor(1:n),
                     variable.name = "topic", 
                     id.vars = "document"))

ggplot(data = viz_df, aes(variable, value, fill = document), ylab = "proportion") +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +  
  coord_flip() +
  facet_wrap(~ document, ncol = n)


```

Here's a neat JSON-based model visualizer, {LDAviz}.  We can use this to visualize the words-on-topics distribution and intertopic distances.  The size of the circles in the LDAvis plot show proportionally the amount of words that belong to each topic, and the space between circles shows the degree to which the circles share words.

```{r LDAvis}
library(LDAvis) #visualization 
library("tsne") #matrix decomposition
svd_tsne <- function(x) tsne(svd(x)$u)
json <- createJSON(
  phi = tmResult$terms, 
  theta = tmResult$topics, 
  doc.length = rowSums(dfm), 
  vocab = colnames(dfm), 
  term.frequency = colSums(dfm),
  mds.method = svd_tsne,
  plot.opts = list(xlab="", ylab="")
)
serVis(json)

```
The relevance param. defines the relationship between the word's topic probability and its lift. Lift is the ratio of the probability of the word in the topic to the probability of the word in the corpus. The more frequent is the word in the corpus, the lower will be its lift. Saliency also gives a idea of both how common the word is in a topic to overall how common it is.  Same vein as tf-idf

